\thetitle

### 2021-05-11 paper/paper.tex
Updates from Overleaf

\begin{abstract}
  % teilw. von IJPR paper uebernommen
  An essential task in manufacturing planning and control is to determine when to release orders to
  the shop floor. One key parameter is the lead time which is the planned time that elapses between
  the release of an order and its completion. Lead times are usually determined based on the
  measured duration orders take to traverse the production system, i.e.\@ the flow times.
  Traditional order release models assume static lead times, although it has been shown that they
  should be set dynamically to be able to react to the dynamic operational characteristics of
  production system. Therefore, we present an order release model which sets lead times dynamically
  by using an reinforcement learning approach. This algorithm is designed especially for
  applications with high variability and periodic feedback to the agent. Therefore, this paper aims
  at providing a proof-of-concept for the proposed algorithm. Therefore, we compare our method on a
  three-stage flow-shop simulation model not only to static order release techniques, but also to
  the commonly used deep Q-Learning algorithm. The results show that both reinforcement learning
  agents have room for improvement for queuing systems with low variability, that our proposed
  method usually outperforms deep Q-Learning and finally for systems with high variability that the
  proposed reinforcement learning algorithm performs very well, outperforming all other approaches.
  %
  % We use a  to compare the algorithm
  % to % and show that our approach outperforms
  % well-known order release mechanisms. We show that in the current version our proposed model using
  % reinforcement learning outperforms some, but not all other tested approaches.
  % EVTL: especially for scenarios with...
  \keywords{operations research, production planning, order release, machine learning, reinforcement
    learning}
\end{abstract}

@@

The first approach by \citet{PaterninaArboleda2001} introduces a predictive order release model by
using reinforcement learning in a single product, four-station serial flow line and compare its
performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP).
%
% Reinforcement learning is an optimisation technique that stems from dynamic programming and its goal
% is to find the best stationary policy for a given problem. This policy is usually provided by
% assessing current and future states (or state-action pairs) of an underlying Markov Decision Process
% (MDP).
% The advantage of reinforcement learning over dynamic programming is that (i) the problem
% space is explored by an agent and thus only expectantly interesting parts of the problem space need
% to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
% as the states are evaluated by consecutively observed states solely.
%
The algorithm of \citet{PaterninaArboleda2001} decides on whether or not to release an order after
each system change, the completion of an operation of any order at any stage or a new order arrival,
and assumes that any unsatisfied demand is lost. Thus, they use a continuous order release method
although in practice order release decisions often need to be made on a periodical basis, e.g.\@
daily (see \citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their tabular
based reinforcement learning agent.
%
Then, \citet{Schneckenreither2019} use several different reinforcement learning algorithms to make
periodic order release decisions for a flow shop production system. The algorithm directly sets lead
times for each product type, which are then used to release the orders in the order pool. They show
that their approach outperforms static order release mechanisms by yielding lower costs, lateness
and standard deviation of lateness, but conclude that research using average reward reinforcement
learning methods harbor optimisation potential for the order release problem.
%
And finally, \citet{schneckenreither2020order} present a flow time estimation procedure to set lead
times dynamically using an artificial neural network, which is used to forecast flow times. By
implementing a rolling horizon order release simulation of the proceeding periods they lift their
approach to a predictive lead time management approach, which is able to detect backorders of future
periods and reacts by releasing orders earlier. Nonetheless, their method is unable to foresee the
triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
prevent the negative effects of the lead time syndrome. Their model outperforms static and reactive
lead time management approaches, especially for scenarios with high utilisation and high variability
in processing times.

@@

In summary, we propose a novel average reward adjusted reinforcement learning algorithm to assess
orders for their release to the shop floor. This is the first average reward adjusted reward
learning algorithm that operates using a artificial neural network as function approximation. To
ensure scalability we adapt a multi-agent approach, where each agent optimises the lead time
management of a single product type. The agents do so by assessing the expected costs for the
possible releases imposed by adapting the lead time. According to the learned estimates each agent
sets a planned lead time for the corresponding product type and with that releases orders into the
production system. The highly optimised value-iteration algorithm uses improved n-step reinforcement
learning with small action-based replay memories and worker agents to infer order release policies
which, especially in case of high variability, outperform not only the static order release
mechanisms, but also the commonly applied deep Q-Learning algorithm~\citep[cf.][]{mnih2015human}.

@@

In case of 80-\(\Exp\) \BILThree{} performs best with costs of \(521.40\), but is not significantly
better than \rlt{1.00} with \(528.54\). Also \rlt{0.99} performs quite well (\(529.00\)). The
service level of all three algorithm are good, but the order is reversed: \rlt{0.99} with
\(87.6\%\), \rlt{1.00} with \(86.6\%\) and \BILThree{} with \(85.8\%\). Again the SFTT of the RL
agents are slightly higher, i.e. \(0.13\)-\(0.20\) periods, than the ones of the \BIL{} release
procedures. The other three \BIL{} methods, \BILOne{}, \BILTwo{} and \BILFour{}, are performing
rather bad (\(1,338.44\), \(679.24\) and \(591.82\) respectively). Interesting to see is that in
this scenario \rlt{0.99} and \rlt{1.00} handle the due date performance better than \qlt{0.99}, as
can be seen in the backorder costs (BOC). Here the former two accumulate costs of \(185.69\) and
\(198.46\), while \qlt{0.99} faces costs of \(225.15\). This results in timing costs (\fgibocName{})
of \(439.16\) for \rlt{0.99} and \(441.37\) for \rlt{1.00}, which also means a better timing cost
than \BILThree{} with \(442.37\). For this scenario the actual measured average reward (average
costs) of about \(75.5\) for both \ARA{} variants and learned average rewards of \(70.53\) for
\rlt{0.99} and \(72.94\) for \rlt{1.00} are a little bit offset. This probably results from the
higher variability in the system.

Finally, for the case of 90-\(\Exp\) \rlt{1.00} performs much better than the other variants with
cots of \(986.32\). It is followed by \rlt{0.99} with costs of \(1,072.06\) and rather far, but not
significantly, afterwards follows \BILFour{} (\(1,104.29\)). As before the shopfloor throughput time
(SFTT) is higher for the RL variants, and with that of course also the work in process costs (WIPC).
The backorder costs of \rlt{1.00} (\(539.80\)) are the lowest of all tested methods. Like for the
overall costs also in terms of timing costs (\fgibocName{}) \rlt{1.00} performs best (\(825.87\)),
followed by \rlt{0.99} (\(914.03\)) and \BILFour{} (\(957.21\)).
%
\BILThree{} and \ql{0.99} perform quite similar in the sense of accumulated costs, \(1,175.88\) and
\(1,179.46\) respectively. But the other measures show that \ql{0.99} has a better timing
performance, thus also a slightly high service level with \(71.5\%\) as compared to \BILThree{} with
\(70.6\%\). The best service level is achieved by \BILFour{} with \(82.5\%\), before \rlt{1.00}
\(79.6\%\), but with a high cost as the difference of SUM is \(117.97\).
%
Interesting to see is that the measured average reward and the learned average reward are very
different for both \ARA{} variants (about \(30\)-\(40\)). This is probably due to the very high
variability in the system, but might also be an indicator that the agents need more time for
learning. However, due to the high computational complexity we encountered we leave this to future
work.

Overall we can see that the RL agents, and especially the established \ARA{} algorithm performs very
well in systems with high variability. In case of the less variable systems the results reveal that
there is still potential for the improvement of reinforcement learning variants.

@@

The results suggest that for high variability the RL learning agents, and here especially the novel
\ARA{} variant, performs very well. In case of medium variability the achieved results are as good
as with the \BIL{} variants. And for low variability one can see that the RL agents still have room
for improvement, as \BIL{} outperforms them. When comparing deep Q-Learning to the established
\ARA{} algorithm, we can see that for all cases, except one (80-\(\Unif\)), \ARA{} performs better.

>>>>

\begin{abstract}
  % teilw. von IJPR paper uebernommen
  An essential task in manufacturing planning and control is to determine when to release orders to
  the shop floor. One key parameter is the lead time which is the planned time that elapses between
  the release of an order and its completion. Lead times are usually determined based on the
  measured duration orders take to traverse the production system, i.e.\@ the flow times.
  Traditional order release models assume static lead times, although it has been shown that they
  should be set dynamically to be able to react to the dynamic operational characteristics of
  production system. Therefore, we present an order release model which sets lead times dynamically
  by using an reinforcement learning approach. This algorithm is designed especially for
  applications with high process variability and periodic feedback to the agent. Therefore, this paper aims
  at providing a proof-of-concept for the proposed algorithm. Therefore, we compare our method on a
  three-stage flow-shop simulation model not only to static order release techniques, but also to
  the commonly used deep Q-Learning algorithm. The results show that both reinforcement learning
  agents have room for improvement for queuing systems with low variability, that our proposed
  method usually outperforms deep Q-Learning and finally for systems with high utilisation that the
  proposed reinforcement learning algorithm performs very well, outperforming all other approaches.
  %
  % We use a  to compare the algorithm
  % to % and show that our approach outperforms
  % well-known order release mechanisms. We show that in the current version our proposed model using
  % reinforcement learning outperforms some, but not all other tested approaches.
  % EVTL: especially for scenarios with...
  \keywords{operations research, production planning, order release, machine learning, reinforcement
    learning}
\end{abstract}

@@

The first approach by \citet{PaterninaArboleda2001} introduces a predictive order release model by
using reinforcement learning in a single product, four-station serial flow line and compare its
performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP).
%
% Reinforcement learning is an optimisation technique that stems from dynamic programming and its goal
% is to find the best stationary policy for a given problem. This policy is usually provided by
% assessing current and future states (or state-action pairs) of an underlying Markov Decision Process
% (MDP).
% The advantage of reinforcement learning over dynamic programming is that (i) the problem
% space is explored by an agent and thus only expectantly interesting parts of the problem space need
% to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
% as the states are evaluated by consecutively observed states solely.
%
The algorithm of \citet{PaterninaArboleda2001} decides on whether or not to release an order after
each system change, the completion of an operation of any order at any stage or a new order arrival,
and assumes that any unsatisfied demand is lost. Thus, they use a continuous order release method
although in practice order release decisions often need to be made on a periodical basis, e.g.\@
daily (see \citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their tabular
based reinforcement learning agent.
%
Then, \citet{Schneckenreither2019} use several different reinforcement learning algorithms to make
periodic order release decisions for a flow shop production system. The algorithm directly sets lead
times for each product type, which are then used to release the orders in the order pool. They show
that their approach outperforms static order release mechanisms by yielding lower costs, lateness
and standard deviation of lateness, but conclude that research using average reward reinforcement
learning methods harbor optimisation potential for the order release problem.
%
And finally, \citet{schneckenreither2020order} present a flow time estimation procedure to set lead
times dynamically using an artificial neural network, which is used to forecast flow times. By
implementing a rolling horizon order release simulation of the proceeding periods they lift their
approach to a predictive lead time management approach, which is able to detect backorders of future
periods and reacts by releasing orders earlier. Nonetheless, their method is unable to foresee the
triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
prevent the negative effects of the lead time syndrome. Their model outperforms static and reactive
lead time management approaches, especially for scenarios with high utilisation.

@@

In summary, we propose a novel average reward adjusted reinforcement learning algorithm to assess
orders for their release to the shop floor. This is the first average reward adjusted reward
learning algorithm that operates using a artificial neural network as function approximation. To
ensure scalability we adapt a multi-agent approach, where each agent optimises the lead time
management of a single product type. The agents do so by assessing the expected costs for the
possible releases imposed by adapting the lead time. According to the learned estimates each agent
sets a planned lead time for the corresponding product type and with that releases orders into the
production system. The highly optimised value-iteration algorithm uses improved n-step reinforcement
learning with small action-based replay memories and worker agents to infer order release policies
which, especially in case of high utilisation, outperform not only the static order release
mechanisms, but also the commonly applied deep Q-Learning algorithm~\citep[cf.][]{mnih2015human}.

@@

In case of 80-\(\Exp\) \BILThree{} performs best with costs of \(521.40\), but is not significantly
better than \rlt{1.00} with \(528.54\). Also \rlt{0.99} performs quite well (\(529.00\)). The
service level of all three algorithm are good, but the order is reversed: \rlt{0.99} with
\(87.6\%\), \rlt{1.00} with \(86.6\%\) and \BILThree{} with \(85.8\%\). Again the SFTT of the RL
agents are slightly higher, i.e. \(0.13\)-\(0.20\) periods, than the ones of the \BIL{} release
procedures. The other three \BIL{} methods, \BILOne{}, \BILTwo{} and \BILFour{}, are performing
rather bad (\(1,338.44\), \(679.24\) and \(591.82\) respectively). Interesting to see is that in
this scenario \rlt{0.99} and \rlt{1.00} handle the due date performance better than \qlt{0.99}, as
can be seen in the backorder costs (BOC). Here the former two accumulate costs of \(185.69\) and
\(198.46\), while \qlt{0.99} faces costs of \(225.15\). This results in timing costs (\fgibocName{})
of \(439.16\) for \rlt{0.99} and \(441.37\) for \rlt{1.00}, which also means a better timing cost
than \BILThree{} with \(442.37\). For this scenario the actual measured average reward (average
costs) of about \(75.5\) for both \ARA{} variants and learned average rewards of \(70.53\) for
\rlt{0.99} and \(72.94\) for \rlt{1.00} are a little bit offset. This probably results from the
higher variability and utilisation in the system.

Finally, for the case of 90-\(\Exp\) \rlt{1.00} performs much better than the other variants with
cots of \(986.32\). It is followed by \rlt{0.99} with costs of \(1,072.06\) and rather far, but not
significantly, afterwards follows \BILFour{} (\(1,104.29\)). As before the shopfloor throughput time
(SFTT) is higher for the RL variants, and with that of course also the work in process costs (WIPC).
The backorder costs of \rlt{1.00} (\(539.80\)) are the lowest of all tested methods. Like for the
overall costs also in terms of timing costs (\fgibocName{}) \rlt{1.00} performs best (\(825.87\)),
followed by \rlt{0.99} (\(914.03\)) and \BILFour{} (\(957.21\)).
%
\BILThree{} and \ql{0.99} perform quite similar in the sense of accumulated costs, \(1,175.88\) and
\(1,179.46\) respectively. But the other measures show that \ql{0.99} has a better timing
performance, thus also a slightly high service level with \(71.5\%\) as compared to \BILThree{} with
\(70.6\%\). The best service level is achieved by \BILFour{} with \(82.5\%\), before \rlt{1.00}
\(79.6\%\), but with a high cost as the difference of SUM is \(117.97\).
%
Interesting to see is that the measured average reward and the learned average reward are very
different for both \ARA{} variants (about \(30\)-\(40\)). This is probably due to the high 
variability and utilisation in the system, but might also be an indicator that the agents need more time for
learning. However, due to the high computational complexity we encountered we leave this to future
work.

Overall we can see that the RL agents, and especially the established \ARA{} algorithm performs very
well in systems with high utilisation. In case of the less utilised systems the results reveal that
there is still potential for the improvement of reinforcement learning variants.

@@

The results suggest that for high utilisation the RL learning agents, and here especially the novel
\ARA{} variant, performs very well. In case of medium utilisation the achieved results are as good
as with the \BIL{} variants. And for low utilisation one can see that the RL agents still have room
for improvement, as \BIL{} outperforms them. When comparing deep Q-Learning to the established
\ARA{} algorithm, we can see that for all cases, except one (80-\(\Unif\)), \ARA{} performs better.
